### 2019-01-17 main.tex
draft

\input{header.tex}

\title{Coupling microsimulation models with land-use transport models}

%\author{\noun{Juste Raimbault}}
\date{}

\maketitle

\justify

\begin{abstract}
	We describe here preliminary investigations towards the coupling of a land-use transport interaction model with a population microsimulation model, at the scale of a country. We situate the approach within the scientific landscape in terms of citation networks around works dealing with microsimulation. We then compare the characteristics of the two models and build on their complementarity to suggest research directions towards this coupling. 
\end{abstract}

%%%%%%%%%%%%
\section{Introduction}

A strong need of multi-scalar models to be applied to sustainable policy making is already shaping the construction of the next generation of territorial models \cite{Rozenblat2018}. Indeed, the coupling of complementary modeling approaches to the same systems is coined by \cite{banos2013pour} as an essential step forward in social simulation to be able to better tackle the complexity of such systems. The modeling activity must furthermore be highly integrated with other knowledge domains \cite{raimbault2017applied}, in particular empirical and theoretical approaches, to foster the production of applicable complex knowledge.

The QUANT land-use transport interaction model \cite{batty2019generalized} responds to this constraints as it integrates fine data for population, employments and transportation at the scale of all UK, while keeping a simple and interpretable structure in the process implemented. Integrating it with a complementary model, possibly at an other scale, would open new possibilities for the design of territorial policies.

In particular, population structure, demographic processes and dynamic aspects are not taken into account in the QUANT model, making microsimulation models \cite{birkin2011spatial} good candidates to be coupled with this model. The purpose of the type of microsimulation models we consider is to provide estimates of spatio-demographic trajectories of territories with a high precision, having an interest in themselves for projections, but also for local policies taking into account spatial heterogeneity and a variety of social processes which can be coupled with microsimulation (economics, crime, retail, health, environmental issues, etc.).

There does not seem to be many works in the literature coupling spatial interaction models with microsimulation models

%%%%%%%%%%%%
\section{Scientific landscape}

For a better grasp of the issues at stake in the approach taken, we propose to map the scientific landscape around works in microsimulation. We apply a part of the method introduced by \cite{raimbault2017exploration}, namely backward citation network reconstruction and mapping, using the open source tools provided. The method to construct the citation network starts from an initial corpus, and retrieves citing references for each paper in the corpus, recursively up to a specified depth. Citation data is crawled from google scholar and results are conditional to this data source (which has a good coverage in comparison to the WoS or Scopus, but includes grey literature and preprints).

We construct the initial corpus with the 200 first responses to the request ``\texttt{microsimulation}'' and construct the citation network at depth 2 from these 200 references. We obtain a network with $\left|V\right|=146774$ nodes and $\left|E\right|=210341$ links, and an average in-degree of 1.43 (3.01 conditionally to have at least one citation). A Louvain community detection on the undirected network yields a directed modularity of 0.69, suggesting strong disciplinary patterns within the network.

It is interesting to remark that the first level (initial corpus) is highly integrated in terms of citations, since it exhibits an average in-degree of 6.46 on the giant component (181 out of 200 references), as shown in Fig.~\ref{fig:citnwdepth0}. The communities obtained already suggest disciplinary and thematic patterns: a large community (purple) includes the SimBritain model \cite{ballas2005simbritain}, issues about validation and synthetic populations, and may be interpreted as a methodological community; an other large community (green) deals with policy application of microsimulation models; and an other one with inequalities (blue); two smaller communities are focused on health applications of microsimulation (dark green) and the link with agent-based models (red), whereas the smallest community is difficult to classify and includes the Sverige model \cite{holm2002sverige}. The papers associated to the model we will study are distributed in the policy community \cite{wu2012moses}, the methodological community \cite{townend2009moses} and the unclassified community \cite{rees2017moses}.

When studying the network at depth 1 (Fig.~\ref{fig:citnwdepth1}), a map of disciplines using microsimulation begin to emerge, and communities are at a lower granularity. Modularity is much higher (0.70). The proper microsimulation papers are split in a large community (blue) and two smaller ones (red and turquoise). We then observe two highly clustered communities corresponding to inequalities (green) and spatial statistics (purple). A community (orange) deals with the links with agent-based modeling. Finally, a community (in yellow) situated at the interface of ABM, microsimulation and spatial statistics deals with transportation and LUTI models, what is of particular interest for our positioning since it confirms the disciplinary relevance of coupling the two approaches and furthermore highlights issues which will be of first importance (the relation to ABMs for example).

At depth 2, the network shown in Fig.~\ref{fig:citnwdepth2} describes the much broader scientific environment of microsimulation models. We still obtain a highly clustered community around inequalities, whereas health geography and LUTI are integrated with spatial statistics and two other communities dealing with two other approaches to inequalities. An other isolated and clustered community focuses on epidemiology. The broad microsimulation clusters acts as a backbone of the network, having roughly a similar proximity with the different domains (which is due to the construction process of the network), at the exception of a high integration with geosimulation and Luti models. This higher integration confirms the relevance of the approach proposed.

Sensitivity to the initial corpus, by changing the number of initial references (and grasping smaller fields of application e.g.) or by starting with a list of specified papers, remains to be checked.

%%%%%%%%%%%%
\begin{figure}
  \includegraphics[width=\linewidth]{figures/microsim_depth0.png}
  \caption{Giant component of the citation network core at depth 0, i.e. 181 references out of the initial corpus with 200 references and the citation links between these. Average in-degree is 6.46 and Louvain modularity is 0.32.}
  \label{fig:citnwdepth0}
\end{figure}
%%%%%%%%%%%%

%%%%%%%%%%%%
\begin{figure}
  \includegraphics[width=\linewidth]{figures/microsim_depth1.png}
  \caption{Citation network core at depth 1. The core is constructed by removing all nodes having no citation, yielding a network of size $\left|V\right|=4846$ nodes and $\left|E\right|=15590$ links, average in-degree of 3.22 and modularity 0.70.}
  \label{fig:citnwdepth1}
\end{figure}
%%%%%%%%%%%%

%%%%%%%%%%%%
\begin{figure}
  \includegraphics[width=\linewidth]{figures/microsim_depth2_corehigher.png}
  \caption{Citation network core at depth 2. For visibility purposes, we show here a ``higher order'' core, obtained by iteratively removing nodes of degree one until having only nodes with at least degree two. We have $\left|V\right|=31294$ and $\left|E\right|=94861$, an average in-degree of 3.03 and a modularity of 0.68.}
  \label{fig:citnwdepth2}
\end{figure}
%%%%%%%%%%%%

%%%%%%%%%%%%
\section{Complementarity of QUANT and MOSES}

We give in Table~\ref{tab:comparison} a rough comparison of QUANT and MOSES models main properties. This highlights the advantages of each and how they can complement each other.

%%%%%%%%%%%%
\begin{table}
\begin{threeparttable}
	\caption{Main characteristics of QUANT and MOSES models.}\label{tab:comparison}
	\centering
	\begin{tabular}{|c|p{6cm}|p{6cm}|}
	\toprule
	& QUANT & MOSES \\
	\midrule
Time scale & 10 years $^{\ast 1}$ & 30 years \\
Spatial scale & UK & UK \\
Spatial resolution & Ward & Ward \\
Static/Dynamic & Equilibrium (static) & Dynamic $^{\ast 2}$ \\
Randomness & Deterministic & Monte-Carlo simulations \\
Process: Transportation & 3 modes $^{\ast 3}$ & NA \\
Process: Economic activities & Accessibility-based relocations & NA \\
Process: Demographics & NA & Data-driven $^{\ast 4}$ \\
Process: Migration flows & Accessibility-based relocations & Data-driven \\
\bottomrule
	\end{tabular}
	\begin{tablenotes}
      \small
      \item $^{\ast 1}$ stylized time scale of metropolitan relocations \cite{wegener2004land}
      \item $^{\ast 2}$ model structure is however stationary - this is an important issue for a possible application to 2070
      \item $^{\ast 3}$ discrete choice for transportation mode is not implemented in QUANT \textbf{(?)} 
      \item $^{\ast 4}$ probabilities of basic demographic events (ageing, mortality, fertility, household formation, health, migration) are computed from data
    \end{tablenotes}
	\end{threeparttable}
\end{table}
%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
%% Biblio
%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{apalike}
\bibliography{biblio}

\end{document}

### 2019-01-18 main.tex
.

@@
 The purpose of the type of microsimulation models we consider is to provide estimates of spatio-demographic trajectories of territories with a high precision, having an interest in themselves for projections, but also for local policies taking into account spatial heterogeneity and a variety of social processes which can be coupled with microsimulation (economics, crime, retail, health, environmental issues, etc.).
+The MOSES model embeds various heterogenous data and is able to simulate demographic trajectories at the Ward level on a 30 years time span \cite{townend2009moses,Wu2013} (this model is distinct from a much older Moses microsimulation model focused on economic processes \cite{eliasson1989moses}).
+A user interface and the connexion to computation infrastructure allow its use for real decision making problems \cite{birkin2009moses}.
+It has been applied for example for the study of the spatial distribution of crime by coupling it with an agent-based model \cite{malleson2012analysis}.
+We will develop here potential directions to couple the QUANT model with the MOSES model (or more precisely its successor, called SPENCER).
 
-There does not seem to be many works in the literature coupling spatial interaction models with microsimulation models
+The literature is rather sparse on studies coupling spatial interaction models with microsimulation models.
+Examples include configurations where the spatial and demographic structure of population plays a role in its interaction with the object of study, such as retail \cite{Nakaya2007}, or in spatial epidemiology.
+Beside possible policy applications, our approach can therefore also provide methodological contributions.
 
@@
 Sensitivity to the initial corpus, by changing the number of initial references (and grasping smaller fields of application e.g.) or by starting with a list of specified papers, remains to be checked.
+Also some parts of literature at the interface of agent-based modeling and microsimulation, which terminology is closer to ABM (e.g.
+the MATSIM model in the case of transportation), may have been missed here.
+A more systematic review as done in \cite{raimbault2018caracterisation} in the case of models of interactions between transportation and territories, combining a broader initial semantic search and a corpus based on a preliminary literature review validated by expert knowledge, should provide a more exhaustive overview of the scientific landscape.
 
@@
 
-We give in Table~\ref{tab:comparison} a rough comparison of QUANT and MOSES models main properties. This highlights the advantages of each and how they can complement each other.
+We give in Table~\ref{tab:comparison} a rough comparison of QUANT and MOSES models main properties.
+This highlights the advantages of each and how they can complement each other.
+First of all, spatial and temporal scales, and spatial resolution, are reasonably comparable, implying that a coupling of these model should naturally be at the same level (although agent granularity is naturally different, the models are for this reason not exactly at the same level in terms of agents).
+The advantage should then be in the coupling of processes, by taking into account complementary aspects of the urban system that each model emphasizes.
+Indeed, when surveying processes included among the different components of the urban system, we observe that transportation and economic processes are not simulated in MOSES whereas they are part of the core of QUANT.
+On the contrary, demographic evolution is not simulated in QUANT but highly precise in MOSES.
+Finally, the movement of population in space is included in both but with different paradigms, since the microsimulation model uses migration probabilities estimated directly from real migration data whereas the Luti model distributes population according to accessibility patterns.
+The gordian knot of coupling will for this reason be on spatial population processes, as other aspects can be included without conflicting with the representation in the other model.
 
@@
 Spatial resolution & Ward & Ward \\
+Agent granularity & Aggregated counts & Person level \\
 Static/Dynamic & Equilibrium (static) & Dynamic $^{\ast 2}$ \\
@@
 
+Possible coupling scenarios, with associated questions and potential outcomes, can be proposed:
+\begin{itemize}
+	\item Weak coupling Luti $\rightarrow$ microsimulation: using sequentially outputs of Quant as inputs of Moses.
+For example, evaluating a new transportation scenario will provide predictions for new populations, employments and commuting flows.
+Distributing the synthetic population and its movements conditionally to this new distribution (respecting reasonably some statistical properties of the original data, what may pose difficult synthetic data generation problems), and running the microsimulation model with this new parametrization, could provide micro-based measures of the impact of the new transportation line (e.g.
+exposure to pollution, access to health facilities, segregation, etc.).
+In that case, it does not seem to make much sense to use the aggregated prediction of the microsimulation model, since the uncertainty on input data will be high as a output of the upstream model (``\textit{garbage in-garbage out}''), and the coupled model has a prospective and scenario-testing function.
+	\item Weak coupling Microsimulation $\rightarrow$ Luti: the other way around, population projections obtained with microsimulation, which should be more precise than relocations by Quant (assuming that the system remains stationary and that exogenous effects are reasonably included), can be used as an input of the Quant model, in order to apply it in future times when for example planning long term infrastructure projects.
+More accurate future populations will provide more accurate generalized accessibility projections given or not some modification of the transportation network.
+The question answered and model function would also be territorial prospective.
+%(\textit{Open question : how do model functions in the sense of the typology of \cite{varenne2017theories} evolve during model coupling, do some typical patterns exist or is it case/scale specific ?
+Given a ``strength'' of models (a partial order on a subset of models given some objectives), are there patterns on the strength of the coupled model ?})
+	\item Both previous proposition make sense assuming large approximations and assumptions.
+However, assuming in the first case that population statistical properties remain similar after urban changes is not accurate, as the composition of population or their behavior may drastically change.
+In the second case, assuming that population projections do not depend much on the changes in the accessibility landscape is also not accurate.
+A strong coupling Microsimulation $\rightarrow$ Luti, in a sense a co-evolutive model between  demographic dynamics and dynamics of accessibility in terms of transportation, population and jobs, would object to these two inaccuracies and in theory be more appropriate.
+It would provide both applications, i.e.
+population exposure impacts of transport and land-use scenarios, but also accurate projections and planning.
+This being said, this approach must be taken with caution.
+Very few co-evolution models of territorial systems (whatever the objects studied) do exist in the literature, and the problem remains relatively opened methodologically, theoretically and empirically (e.g.
+\cite{raimbault2018modeling} extended the city system model of \cite{raimbault2018indirect} to provide a co-evolution model between transportation networks and territories at the macro-scale, which although very novel as such kind of models remains very simple and highly limited).
+For example, one can use Quant population dynamics to constraint microsimulation probabilities at fixed time steps, or one can on the contrary use microsimulation outputs as additional exogenous constraints for Quant which will be useful to redistribute employments and a given transportation scenario on a long time period.
+Many different coupling choices must be more thoroughly listed, studied and possibly compared.
+In that sense, building a strongly coupled model is similar to constructing a new model.
+\end{itemize}
+
+% \section{Other directions}
+% microsim models with other ontologies, e.g. Matsim for transportation ?
+
 %%%%%%%%%%%%%%%%%%%%

### 2019-01-18 main.tex
matsim
Update on Overleaf.

@@
 \input{header.tex}
+\begin{document}
 
@@
 
-%\author{\noun{Juste Raimbault}}
+%\author{}
 \date{}
@@
 
-The QUANT land-use transport interaction model \cite{batty2019generalized} responds to this constraints as it integrates fine data for population, employments and transportation at the scale of all UK, while keeping a simple and interpretable structure in the process implemented.
+The QUANT land-use transport interaction model \cite{batty2019generalized} responds to this constraints as it integrates fine data for population, employments and transportation at the scale of all UK, while keeping a simple structure for the processes implemented.
 Integrating it with a complementary model, possibly at an other scale, would open new possibilities for the design of territorial policies.
@@
 The literature is rather sparse on studies coupling spatial interaction models with microsimulation models.
-Examples include configurations where the spatial and demographic structure of population plays a role in its interaction with the object of study, such as retail \cite{Nakaya2007}, or in spatial epidemiology.
+Examples include configurations where the spatial and demographic structure of population plays a role in its interaction with the object of study, such as retail \cite{Nakaya2007}, in spatial epidemiology \cite{morrissey2010examining}, or for socio-spatial simulations of a particular population \cite{wu2008spatial}.
 Beside possible policy applications, our approach can therefore also provide methodological contributions.
@@
 
-% \section{Other directions}
+\section{Other perspectives}
+
 % microsim models with other ontologies, e.g. Matsim for transportation ?
+
+\begin{itemize}
+	\item Matsim is an agent-based approach to transportation demand which is activity-based \cite{horni2016multi}.
+It can simulate agents behavior within the transportation system (traffic flows) and behavioral patterns generating the travel demand.
+According to \cite{wise2016transportation}, accurate description of transportation remains a point to be explored for agent-based model.
+A weak coupling of Quant and Matsim would be interesting in that sense, and allow to precisely explore the consequence of transportation scenarios.
+	\item More generally, the fuzzy boundary between microsimulation and agent-based modeling may be an interesting point to be investigated, both methodologically and theoretically.
+\end{itemize}
 

### 2019-01-18 main.tex
Update on Overleaf.

@@
 
-A strong need of multi-scalar models to be applied to sustainable policy making is already shaping the construction of the next generation of territorial models \cite{Rozenblat2018}.
+A strong need of multi-scalar models to be applied to sustainable urban policy making is already shaping the construction of the next generation of territorial models \cite{Rozenblat2018}.
 Indeed, the coupling of complementary modeling approaches to the same systems is coined by \cite{banos2013pour} as an essential step forward in social simulation to be able to better tackle the complexity of such systems.

### 2019-02-05 main.tex
Update on Overleaf.

@@
 
+\subsection{Microsimulation models}
+
+\cite{orcutt1957new} introduced the seminal idea of dealing with individual units which led to the emergence of microsimulation approaches in for socio-economic systems.
+
+\cite{birkin2012review} review :
+\begin{itemize}
+    \item Privileged fields of application for policy evaluation: taxation scheme evaluation; pension management anticipation; health (environment-dependant conditions and spatial distribution of health services); transport (traffic microsimulations - \textit{even closer to ABMs in this field, the difference would be the application to policies or network interventions, rather than a principal focus on agent behavior}).
+    \item Static vs Dynamic: static has fixed rules and updates weights e.g, dynamic simulates interactions and endogenous changes; static is cheap computationally.
+    \item Spatial aspect: inclusion of precise local context; various sources and application to projections.
+    \item Need for new hybrid approaches tackling: scale of datasets needed and deployment; no bottom-up feedback in MSMs (i.e.
+long time scale social dynamics); less capabilities in behavioral modeling; problems with validation (\textit{rq: I think it is roughly the same for ABMs and CAs, validation standards are rather sparse}).
+Proposal: space as the bridge between ABM and MSM; GIS as support for ABMs; synthesis of the three.
+(multi-scale / endogeneity etc.
+/ co-evolution for transport - see \cite{batty2019causality} : many (coupled) models, different perspectives).
+Behavioral changes linked to policies and reciprocally can better be taken into account.
+\end{itemize}
+
+\subsection{Differences between microsimulation and agent-based models}
+
+\textbf{Possible main question for the paper: are ABMs and MSMs so different ?} (what about other models ?)
+
+\subsection{Description of the MOSES model}
+
 %%%%%%%%%%%%
-\section{Scientific landscape}
+\section{Mapping of the scientific landscape}
+
+\paragraph{Citation network analysis}
+
+\paragraph{Methods}
 
@@
 Citation data is crawled from google scholar and results are conditional to this data source (which has a good coverage in comparison to the WoS or Scopus, but includes grey literature and preprints).
+
+\subsection{Preliminary results}
 
@@
 
+\textit{Specificity of community detection in citation networks ?
+} \cite{chen2010community} directed top-down modularity maximization (citation communities in PR publications) ; \cite{malliaros2013clustering} review of methods for directed modularity detection.
+
+%%%%%%%%%%%
+\subsection{Interdisciplinarity levels and bridge publications}
+
+Questions to be answered using centrality indices (test specific measures ? \cite{latora2007measure}) and interdisciplinarity measures \cite{leydesdorff2011indicators}
+
 %%%%%%%%%%%%
 \section{Complementarity of QUANT and MOSES}
+
+\textit{Possible meta-analysis ?} $\rightarrow$ collect characteristics for more models; link to disciplinary/semantic characateristics
 
@@
 Spatial scale & UK & UK \\
-Spatial resolution & Ward & Ward \\
-Agent granularity & Aggregated counts & Person level \\
+Spatial resolution & MSOA & Ward \\
+Agent granularity & Aggregated counts & individual level \\
 Static/Dynamic & Equilibrium (static) & Dynamic $^{\ast 2}$ \\
@@
 
-\begin{itemize}
-	\item Matsim is an agent-based approach to transportation demand which is activity-based \cite{horni2016multi}.
+\subsection{Microsimulation and agent-based modeling}
+
+\subsubsection{The MATSim model}
+
+ Matsim is an agent-based approach to transportation demand which is activity-based \cite{horni2016multi}.
 It can simulate agents behavior within the transportation system (traffic flows) and behavioral patterns generating the travel demand.
-According to \cite{wise2016transportation}, accurate description of transportation remains a point to be explored for agent-based model.
+According to \cite{wise2016transportation}, accurate description of transportation remains a point to be explored for agent-based models.
 A weak coupling of Quant and Matsim would be interesting in that sense, and allow to precisely explore the consequence of transportation scenarios.
-	\item More generally, the fuzzy boundary between microsimulation and agent-based modeling may be an interesting point to be investigated, both methodologically and theoretically.
-\end{itemize}
+
+\begin{table}
+\caption{Characteristics of the MATSim model}
+\begin{tabular}{|c|p{6cm}|}
+	\toprule
+	& MATSim \\
+	\midrule
+Time scale & 1 day \\
+Spatial scale & Any scale (mostly metropolitan area)\\
+Spatial resolution & Activities coordinates \\
+Agent granularity & Individual level \\
+Static/Dynamic & Dynamic Stochastic User Equilibrium \\
+Randomness & Stochastic co-evolutionary algorithm (agent level) \\
+Process: Transportation & Several modes; microsimulated\\
+Process: Economic activities & Exogenous \\
+Process: Demographics & NA (agent schedules as input) \\
+Process: Migration flows & NA \textbf{check if origins can change} \\
+Process: Transportation congestion & Link queuing
+\bottomrule
+	\end{tabular}
+\end{table}
+
+ More generally, the fuzzy boundary between microsimulation and agent-based modeling may be an interesting point to be investigated, both methodologically and theoretically.
 

### 2019-03-07 abstract.tex
abstract

\input{header.tex}
\begin{document}

\title{A citation network analysis of a scientific landscape in urban modeling\\\medskip
\textit{Abstract proposal}\medskip\\
\textit{Applied Network Science}\\
\textit{Special Issue Complexity and the City}
}
\author{\noun{Juste Raimbault}$^{1,2,3}$\medskip\\
$^1$ UPS CNRS 3611 ISC-PIF\\
$^2$ CASA, UCL\\
$^3$ UMR CNRS 8504 Géographie-cités
}
%\date{Novembre 2016}
\date{}

\maketitle

\justify

%\begin{abstract}
%\end{abstract}

\pagenumbering{gobble}

%\vspace{-1cm}

\textbf{Keywords: }\textit{Citation network; Interdisciplinarity; Urban Modeling; Land-use Transport Interaction Models; Microsimulation Models}

\bigskip

Several modeling approaches to the evolution of urban and territorial environments have been introduced in the literature, providing complementary viewpoints and methods. This paper contributes to the mapping of this scientific landscape by a citation network analysis. Taking the case study of two landmark types of operational models, namely microsimulation models and land-use transport interaction model, we construct an initial corpus through keyword requests. The backward citation network is then reconstructed up to depth two, yielding a network with $1.47\cdot 10^5$ nodes and $2.10\cdot 10^5$ citation links. Community detection in the network provides endogenous disciplines and approaches, including a broader scope than the initial corpus, with for example spatial statistics and agent-based modeling. We construct measures of interdisciplinarity to identify bridge papers, and investigate possible missing connections. We illustrate these more particularly with the interface of microsimulation, agent-based modeling and land-use transport models, and discuss potential coupling of models filling this gap. We show therein how potential research directions can be unveiled through citation network analysis and provide partial maps of science in the field of urban modeling, which can be used by researchers and practitioners to better contextualize their modeling approaches.

%%%%%%%%%%%%%%%%%%%%
%% Biblio
%%%%%%%%%%%%%%%%%%%%

%\footnotesize

%\bibliographystyle{apalike}
%\bibliography{biblio}

\end{document}

### 2019-03-07 paper.tex
Update on Overleaf.

\input{header.tex}
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

\title{Citation network analysis of a scientific landscape in urban modeling}

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={juste.raimbault@polytechnique.edu}   % email address
]{\inits{JR}\fnm{Juste} \snm{Raimbault}}
\author[
   addressref={aff1,aff2},
   email={john.RS.Smith@cambridge.co.uk}
]{\inits{JRS}\fnm{John RS} \snm{Smith}}

\address[id=aff1]{%                           % unique id
  \orgname{Department of Zoology, Cambridge}, % university, etc
  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgname{CASA, UCL},
  \street{D\"{u}sternbrooker Weg 20},
  \postcode{24105}
  \city{London},
  \cny{UK}
}

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

\begin{abstractbox}

\begin{abstract} % abstract
%\parttitle{First part title} %if any
Several modeling approaches to the evolution of urban and territorial environments have been introduced in the literature, providing complementary viewpoints and methods. This paper contributes to the mapping of this scientific landscape by a citation network analysis. Taking the case study of two landmark types of operational models, namely microsimulation models and land-use transport interaction model, we construct an initial corpus through keyword requests. The backward citation network is then reconstructed up to depth two, yielding a network with $1.47\cdot 10^5$ nodes and $2.10\cdot 10^5$ citation links. Community detection in the network provides endogenous disciplines and approaches, including a broader scope than the initial corpus, with for example spatial statistics and agent-based modeling. We construct measures of interdisciplinarity to identify bridge papers, and investigate possible missing connections. We illustrate these more particularly with the interface of microsimulation, agent-based modeling and land-use transport models, and discuss potential coupling of models filling this gap. We show therein how potential research directions can be unveiled through citation network analysis and provide partial maps of science in the field of urban modeling, which can be used by researchers and practitioners to better contextualize their modeling approaches.
\end{abstract}

\begin{keyword}
\kwd{citation network}
\kwd{interdisciplinarity}
\kwd{urban modeling}
\kwd{land-use transport interaction models}
\kwd{microsimulation models}
\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods and data}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

\subsection*{Community detection}

\subsection*{Measuring interdisciplinarity}

\subsection*{A case study: LUTI and microsimulation}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

%\section*{Competing interests}
%  The authors declare that they have no competing interests.

%\section*{Author's contributions}
%    Text for this section \ldots

%\section*{Acknowledgements}
%  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{biblio}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%  \begin{figure}[h!]
%  \caption{\csentence{Sample figure title.}
%      A short description of the figure content
%      should go here.}
%      \end{figure}

%\begin{figure}[h!]
%  \caption{\csentence{Sample figure title.}
%      Figure legend text.}
%      \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{table}[h!]
%\caption{Sample table title. This is where the description of the table should go.}
%      \begin{tabular}{cccc}
%        \hline
%           & B1  &B2   & B3\\ \hline
%        A1 & 0.1 & 0.2 & 0.3\\
%        A2 & ... & ..  & .\\
%        A3 & ..  & .   & .\\ \hline
%      \end{tabular}
%\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section*{Additional Files}
%  \subsection*{Additional file 1 --- Sample additional file title}
%    Additional file descriptions text (including details of how to   view the file, if it is in a non-standard format or the file extension).  This might   refer to a multi-page table or a figure.
% \subsection*{Additional file 2 --- Sample additional file title}
%    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2019-05-16 paper.tex
Update on Overleaf.

@@
 
-\title{Citation network analysis of a scientific landscape in urban modeling}
+\title{Citation network analysis of urban modeling research}
 
 \author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   addressref={aff1,aff2,aff3},                   % id's of addresses, e.g. {aff1,aff2}
    corref={aff1},                       % id of corresponding address, if any
@@
 \author[
-   addressref={aff1,aff2},
-   email={john.RS.Smith@cambridge.co.uk}
-]{\inits{JRS}\fnm{John RS} \snm{Smith}}
+   addressref={aff2},
+   email={m.vanhoof@ucl.ac.uk}
+]{\inits{MV}\fnm{Maarten} \snm{Vanhoof}}
+\author[
+   addressref={aff2},
+   email={m.batty@ucl.ac.uk}
+]{\inits{MB}\fnm{Michael} \snm{Batty}}
 
 \address[id=aff1]{%                           % unique id
-  \orgname{Department of Zoology, Cambridge}, % university, etc
-  \street{Waterloo Road},                     %
-  %\postcode{}                                % post or zip code
-  \city{London},                              % city
-  \cny{UK}                                    % country
+  \orgname{UPS CNRS 3611 ISC-PIF}, % university, etc
+  %\street{113 rue Nationale},                     %
+  %\postcode{75013}                                % post or zip code
+  \city{Paris},                              % city
+  \cny{France}                                    % country
 }
@@
   \orgname{CASA, UCL},
-  \street{D\"{u}sternbrooker Weg 20},
-  \postcode{24105}
+  %\street{},
+  %\postcode{}
   \city{London},
   \cny{UK}
+}
+\address[id=aff3]{%
+  \orgname{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s},
+  %\street{},
+  %\postcode{}
+  \city{Paris},
+  \cny{France}
 }
@@
 Taking the case study of two landmark types of operational models, namely microsimulation models and land-use transport interaction model, we construct an initial corpus through keyword requests.
-The backward citation network is then reconstructed up to depth two, yielding a network with $1.47\cdot 10^5$ nodes and $2.10\cdot 10^5$ citation links.
+The backward citation network is then reconstructed up to depth two, yielding a network with $4.54\cdot 10^5$ nodes and $8.07\cdot 10^5$ citation links.
 Community detection in the network provides endogenous disciplines and approaches, including a broader scope than the initial corpus, with for example spatial statistics and agent-based modeling.
@@
 
+Urban modeling, defined broadly as quantitative modeling and simulation aiming at understanding urban processes, has always been a highly multidisciplinary and interdisciplinary subject.
+Indeed, several disciplines ranging from Urban Economics~\cite{krugman1998s} to Quantitative Urban Geography~\cite{pumain2017geography} or Urban Planning~\cite{knowles2016sir} provide complementary viewpoints by studying different dimensions of urban phenomena.
+Despite sometimes converging issues, bridges remain rather rare, such as integrating approaches such as the Science of Cities introduced by \cite{batty2013new}.
+One crucial question in that context is to what extent these different approaches to modeling urban issues may be complementary, and what may the potentially fertile couplings in terms of producing new knowledge.
+
+While problem-specific entries, i.e.
+studying a specific system and correspondences between some of its models, will shed light on thematic aspects of this question, one can also consider the knowledge landscape in itself to identify existing approaches and potential links.
+This paper investigates urban modeling at this second level, and proposes to build a mapping of some approaches in urban modeling through citation network analysis.
+Quantitative studies of science have a long history such as in sociology of science~\cite{barnes1982science}, but recent works have witnessed significant progress driven by the emergence of large dataset and digital libraries.
+Maps of science have been constructed \cite{boyack2005mapping} and a ``Science of Science'' would be appearing as an interdisciplinary field~\cite{fortunato2018science}.
+We situate here within that stream of research, using network analysis to better understand the structure of knowledge around the topic of urban modeling.
+
+Citation network has to the best of our knowledge never been systematically applied to the subject of urban modeling, in the sense of a broad and general coverage and large dataset.
+Specific urban issues were however studied through citation network analysis.
+\cite{van2016pacifying} map the debate around urban region polycentricity.
+\cite{raimbault2019exploration} develops a multilayer analysis, coupling citations with semantic content, of the neighborhood of a geography journal, including clusters tackling urban questions.
+
+We construct a large corpus of papers, linked through backward citations.
+Our contribution relies on the following points: 
+% (i) sensitivity analysis - info to reconstruct discipline with such empirical data ?
+% (ii) some methodological insights depending on what we do ?
+% (iii) results on urban modeling in themselves <- Maybe insist only on that ?
+
+The rest of the paper is organized as follows. We first describe the construction of the dataset. We then explore it answering different thematic questions; ..
+
 %%%%%%%%%%%%%%%%%%%%%%
-\section*{Methods and data}
+\section*{Dataset construction}
+
+\subsection*{Citation network collection}
+
+We propose a double entry to build a broad corpus, on complementary but distinct approaches.
+This will ensure first that the collected data is representative due to the corpus size limitations, but secondly that potential couplings are included.
+% REFORMULATE
+
+The initial keyword requests we use are the following: ``microsimulation'', ``microsimulation model'', ``spatial microsimulation model'', ``urban microsimulation model'', ``transport microsimulation model'', ``land use transport interaction model''.
+
+Our corpus is surely not exhaustive to urban modeling as we focused on the aforementioned two main entries, and other requests such as ``urban geosimulation'', ``urban growth model'', ``urban agent based model'' would also be relevant.
+We expect for example, following \cite{wise2016transportation}, to have a strong overlap with our corpus and a potential corpus related to agent-based modeling.
+%urban agent based model, spatial interaction model, spatial interaction agent based model, geosimulation, urban geosimulation, urban cellular automata, urban growth model, urban land use change model, urban land use model
+
+Language statistics (detected using titles). We do not any semantic analysis here and thus do not filter on language.
+% possible additional analysis by isolating language subnetworks - checking modularity of language communities ?.
+Typically: Chinese article are only cited by other Chinese articles.
+similar but less strong effect with French
+
+% main languages
+%    Chinese      Czech      Dutch    English     French   Galician     German Indonesian 
+%     24211        589        872     450149       6425        663       8467        880 
+%   Italian   Japanese     Korean     Polish Portuguese    Russian    Spanish    Swedish 
+%      1641       2457       2080       2284       6348       1981       9234        654 
+%   Turkish 
+%      1196
+
+%%%%%
+% full network
+%IGRAPH 9ceae4d DN-- 467608 834253 -- 
+%+ attr: name (v/c), title (v/c), year (v/c), depth (v/n), priority (v/n), horizontalDepth (v/c), citingFilled (v/c), lang (v/c)
+
+% components(citation)$csize
+% 466610    517      5    246     53     57      6     58      3     10      7      5      7     15      3      2     2      2
+%  2nd comp 517 -> breast cancer ! (should come from "microsimulation" req alone
+% 4th component : agriculture / soils
+
+We keep the giant clusters %TODO cluster statistics
+and the ``core'' of the network, which we define as
 
@@
 
-\subsection*{Community detection}
+\subsection*{Sensitivity analysis}
 
-\subsection*{Measuring interdisciplinarity}
+check modularity = f(horizontal and vertical depth) : insight into the relevant number of initially requested papers
 
-\subsection*{A case study: LUTI and microsimulation}
+Sensitivity of 'width' of paper. Width=the amount of hits in google scholar we use
+
+Sensitivity of the used key words
+
+Overlap for the different keywords combinations, maybe filter papers that are present in all keyword searches
+
+\subsection*{Endogenous disciplines}
+
+% cluster sizes
+
+% Investigate the occurrence of some terms and their membership towards different clusters
+
+%Measuring interdisciplinarity
+
+%The existence of clusters related to different tools or techniques (e.g. Matsim)
+
+%Can we reconstruct the cluster based on iterative search terms in google scholar? -- see where the convergence point is for certain terms
+
+\subsection*{Temporal analysis of communities}
+% Investigate the temporality of the nodes and links % TODO Maarten
+
+% Percolation of links based on frequency of citations or maybe on time also -> is percolation useful at this point ?
+
+%\subsection*{A case study: LUTI and microsimulation}
 
@@
 \section*{Discussion}
+
+% possible developments:
+% with semantic content : 
+%Text mining of abstracts to automatically denote -content- of the clusters\
+% Investigate the location of certain authors\\
+% Investigate the propagation of some of the most important works\\
+% Investigate the content of 'connecting' papers
+
+% TODO : possible subject for Tongji workshop (search chinese kws)
 

### 2019-06-05 paper.tex
Update on Overleaf.
.

@@
 
+% \cite{ramadier2004transdisciplinarity} transdisciplinarity in urban studies
+% \cite{batty2019map} the digital twin - need of reflexivity, hence understanding modeling landscape is crucial
+
 While problem-specific entries, i.e.
@@
 \cite{raimbault2019exploration} develops a multilayer analysis, coupling citations with semantic content, of the neighborhood of a geography journal, including clusters tackling urban questions.
+
+% \cite{giatsidis2019rooted} new measures based on subgraph density for citation networks
 
@@
 
-check modularity = f(horizontal and vertical depth) : insight into the relevant number of initially requested papers
+% check modularity = f(horizontal and vertical depth) : insight into the relevant number of initially requested papers ; Sensitivity of 'width' of paper.
+Width=the amount of hits in google scholar we use ; Sensitivity of the used key words => all in same analysis
 
-Sensitivity of 'width' of paper. Width=the amount of hits in google scholar we use
-
-Sensitivity of the used key words
+Before being able to proceed to any thematically relevant network analysis, a sensitivity analysis of network properties as a function of construction parameters is necessary.
+More particularly, the results may be strongly dependent on the number of search results retained and on the requests used.
+This preliminary aims at checking this dependency and extract the most robust subnetwork possible for the rest of the analysis.
 

### 2019-06-05 paper.tex
Update on Overleaf.

@@
 
-Urban modeling, defined broadly as quantitative modeling and simulation aiming at understanding urban processes, has always been a highly multidisciplinary and interdisciplinary subject.
+Urban modeling, defined broadly as quantitative modeling and simulation aiming at understanding urban processes, has always been a highly multidisciplinary and interdisciplinary subject \cite{ramadier2004transdisciplinarity}.
 Indeed, several disciplines ranging from Urban Economics~\cite{krugman1998s} to Quantitative Urban Geography~\cite{pumain2017geography} or Urban Planning~\cite{knowles2016sir} provide complementary viewpoints by studying different dimensions of urban phenomena.
@@
 One crucial question in that context is to what extent these different approaches to modeling urban issues may be complementary, and what may the potentially fertile couplings in terms of producing new knowledge.
-
-% \cite{ramadier2004transdisciplinarity} transdisciplinarity in urban studies
-% \cite{batty2019map} the digital twin - need of reflexivity, hence understanding modeling landscape is crucial
+Indeed, a good knowledge of the scientific landscape, or in other terms a certain level of reflexivity, is crucial for urban modeling studies since as \cite{batty2019map} points out, models become digital twins of modelled systems which themselves play a role in the studied system.
 
@@
 
-% \cite{giatsidis2019rooted} new measures based on subgraph density for citation networks
+%\cite{giatsidis2019rooted} propose  new measures based on subgraph density for citation networks
 
 We construct a large corpus of papers, linked through backward citations.
-Our contribution relies on the following points: 
+Our contribution relies on the following points: (i) the open dataset we make available can be of significant interest for researcher interested in urban modeling; (ii) we map a broad part of approaches to urban modeling, and more particularly clarify the relations between microsimulation, land-use transport modeling, and agent-nased modeling.
 % (i) sensitivity analysis - info to reconstruct discipline with such empirical data ?
@@
 
-The rest of the paper is organized as follows. We first describe the construction of the dataset. We then explore it answering different thematic questions; ..
+The rest of the paper is organized as follows.
+We first describe the construction of the dataset, i.e.
+the initial assumptions and data collection process, and the filtering steps to work on a relevant network.
+We then study its sensitivity to construction parameters, what allows to comment on endogenous disciplines and study the evolution of communities in time.
+We finally discuss the implication of these results, for example in the context of coupling approaches.
 
@@
 
-\subsection*{Citation network collection}
+\subsection*{Initial seed}
 
-We propose a double entry to build a broad corpus, on complementary but distinct approaches.
+We propose a multiple keyword entry to build a broad corpus, on complementary but distinct approaches.
 This will ensure first that the collected data is representative due to the corpus size limitations, but secondly that potential couplings are included.
@@
 We expect for example, following \cite{wise2016transportation}, to have a strong overlap with our corpus and a potential corpus related to agent-based modeling.
+Similarly, other sub-fields of urban modeling such as cellular automata \cite{white1993cellular} or geosimulation \cite{benenson2004geosimulation} are expected to be highly interwined with our corpus from a thematic viewpoint.
 %urban agent based model, spatial interaction model, spatial interaction agent based model, geosimulation, urban geosimulation, urban cellular automata, urban growth model, urban land use change model, urban land use model
 
-Language statistics (detected using titles). We do not any semantic analysis here and thus do not filter on language.
-% possible additional analysis by isolating language subnetworks - checking modularity of language communities ?.
-Typically: Chinese article are only cited by other Chinese articles.
-similar but less strong effect with French
-
-% main languages
-%    Chinese      Czech      Dutch    English     French   Galician     German Indonesian 
-%     24211        589        872     450149       6425        663       8467        880 
-%   Italian   Japanese     Korean     Polish Portuguese    Russian    Spanish    Swedish 
-%      1641       2457       2080       2284       6348       1981       9234        654 
-%   Turkish 
-%      1196
+\subsection*{Citation network collection}
 
@@
 
+\subsection*{Network filtering}
+
+Language statistics (detected using titles).
+We do not include in our analysis any semantic analysis and thus do not filter on language of papers.
+Note that as we work on citation networks, cross-language citations are a common feature and different languages will not produce isolated clusters.
+In any case, a language detection with the python package \texttt{polyglot} confirms that most of the corpus is in English (86.6\%), secondary main language being Mandarin (4.7\%), Spanish (1.8\%), German (1.6\%), French (1.2\%) and Portuguese (1.2\%).
+% possible additional analysis by isolating language subnetworks - checking modularity of language communities ?.
+Typically: Chinese article are only cited by other Chinese articles.
+similar but less strong effect with French
+
+%     Czech    Swedish   Galician      Dutch Indonesian    Turkish   
+% 0.1129035  0.1259161  0.1274470  0.1672503  0.1693553  0.2309738  
+% Italian    Russian     Korean     Polish 
+%0.3155558  0.3832979  0.3987982  0.4384102 
+%  Japanese Portuguese     French     German    Spanish  
+% 0.4720899  1.2195495  1.2339017  1.6231318  1.7783263 
+% Chinese    English 
+%4.6506688 86.5524236
+
 % components(citation)$csize
@@
 
-We keep the giant clusters %TODO cluster statistics
-and the ``core'' of the network, which we define as
+We keep the giant connected cluster of the resulting network, which covers 99.8\% of all nodes (yielding at this step $\left|V\right| = 466,610$ nodes).
+Inspection of removed clusters confirms that these are not relevant to urban issues (medical or chemistry microsimulations for example).
+Because of the way citation networks are organized and how our dataset was constructed, the full network contains a large part of ``graps'' which consist in several uncited papers citing the same one.
+We choose thus to work on the ``core'' of the network, which we construct by successively removing nodes of degree less or equal to one until there are no more such nodes.
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section*{Results}
+\section*{Mapping the scientific landscape}
 
@@
 
-Overlap for the different keywords combinations, maybe filter papers that are present in all keyword searches
+We first investigate how network topology varies as a function of horizontal depth.
 
-\subsection*{Endogenous disciplines}
+%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{}
+    \label{fig:modularity-hdepth}
+\end{figure}
+%%%%%%%%
+
+We then assess that a certain level of overlap exists between the sub-networks originating from different initial requests. As shown in Fig.~\ref{fig:coverage-hdepth}
+
+%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics{figures/coveragesubnws.png}
+    \caption{}
+    \label{fig:coverage-hdepth}
+\end{figure}
+%%%%%%%%
+
+\subsection*{Reconstructing endogenous disciplines}
 
 % cluster sizes
+The distribution of community size is fat-tailed in a power-law manner
 
@@
 %The existence of clusters related to different tools or techniques (e.g. Matsim)
-
-%Can we reconstruct the cluster based on iterative search terms in google scholar? -- see where the convergence point is for certain terms
 
@@
 % Investigate the content of 'connecting' papers
+
+%Can we reconstruct the cluster based on iterative search terms in google scholar? -- see where the convergence point is for certain terms
 

### 2019-06-11 paper.tex
Update on Overleaf.
.

@@
 We expect for example, following \cite{wise2016transportation}, to have a strong overlap with our corpus and a potential corpus related to agent-based modeling.
-Similarly, other sub-fields of urban modeling such as cellular automata \cite{white1993cellular} or geosimulation \cite{benenson2004geosimulation} are expected to be highly interwined with our corpus from a thematic viewpoint.
+Similarly, other sub-fields of urban modeling such as cellular automata \cite{white1993cellular} or geosimulation \cite{benenson2004geosimulation} are expected to be highly intertwined with our corpus from a thematic viewpoint.
+To what extent these are indeed present in the corpus is an important property we will investigate.
 %urban agent based model, spatial interaction model, spatial interaction agent based model, geosimulation, urban geosimulation, urban cellular automata, urban growth model, urban land use change model, urban land use model
@@
 \subsection*{Citation network collection}
+
+To collect bibliographic data, we follow the method introduced by \cite{raimbault2019exploration} which gathers a corpus from google scholar, and provide the open tools to do so.
+This source is not the best regarding metadata quality for example, but is much more exhaustive on grey literature or preprints and thus reflects the actual state of a field as it is practiced rather than a filtered and lagged in time image of it.
+% remark : for some discplines / domains, this should be significantly important - I'm thinking of physics with preprints or computer science which publishes mostly in proceedings.
+
+% TODO here introduce the horizontal depth parameter
+The initial keyword requests are achieved given a parameter $h_d$ (horizontal depth), corresponding to the number of papers kept for each request.
+Intuitively, one could expect to retain core papers to the subject for low $h_d$ values while covering broader landscapes for high values.
+This will be studied with a sensitivity analysis below.
 
@@
 %+ attr: name (v/c), title (v/c), year (v/c), depth (v/n), priority (v/n), horizontalDepth (v/c), citingFilled (v/c), lang (v/c)
+
+The full network has a count of $\left|V\right|= 467,608$ nodes and $\left|E\right|=834,253$ citation links.
 
@@
 
-We keep the giant connected cluster of the resulting network, which covers 99.8\% of all nodes (yielding at this step $\left|V\right| = 466,610$ nodes).
+We keep the giant connected cluster of the resulting network, which covers 99.8\% of all nodes.
+%(yielding at this step $\left|V\right| = 466,610$ nodes).
  Inspection of removed clusters confirms that these are not relevant to urban issues (medical or chemistry microsimulations for example).
@@
 We choose thus to work on the ``core'' of the network, which we construct by successively removing nodes of degree less or equal to one until there are no more such nodes.
+We finally remove manually two communities corresponding to medical science and economics of taxes papers, which appear to be connected (most probably through methodological papers) but are not relevant for our study.
+The final filtered network has a size of % TODO
+  .
 
@@
     \centering
-    \includegraphics{}
+    %\includegraphics{figures/}}}
     \caption{}
@@
     \includegraphics{figures/coveragesubnws.png}
-    \caption{}
+    \caption{Relative coverage of sub-networks for each initial request, as a function of horizontal depth $h_d$.}
     \label{fig:coverage-hdepth}
@@
 
+% rq ; microsim does not appear explicitely in main title -> finally "in minority" in urban modeling ;  other terms (in case of same approach) and other approaches are preefered.
+
 % cluster sizes
-The distribution of community size is fat-tailed in a power-law manner
+The distribution of community size is fat-tailed in a power-law manner, as scientific disciplines generally are. % TODO ref on that ?
+ We can thus limit our study to the most important communities.
 
 % Investigate the occurrence of some terms and their membership towards different clusters
+
+% graph visualization and relative positioning of disciplines
+
+We show in Fig.~\ref{fig:networkviz} a visualization of the network obtained with $h_d=3$.
+
+% Q : what about older nws ? also show interesting entry through one kw only. - we can not visualize everything unfortunately.
+%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics{figures/corehdepth3filteredLOWRES.png}
+    \caption{Network visualization for $h_d = 3$. Node color give community obtained with the Louvain algorithm.}
+    \label{fig:networkviz}
+\end{figure}
+%%%%%%%%
 
@@
 
+\subsubsection*{A case study: LUTI, microsimulation and agent-based modeling}
+
+% specific viz including 4 "core thematic clusters" -> abm, microsim, luti + complex systems (glue ?) [no transport abm : microsim]
+% - one hypothesis (seen in temporal analysis) : complex system/complex network approaches as linking prviously less related approaches ?
+
 \subsection*{Temporal analysis of communities}
@@
 % Percolation of links based on frequency of citations or maybe on time also -> is percolation useful at this point ?
-
-%\subsection*{A case study: LUTI and microsimulation}
 

### 2019-06-11 wp.tex
Update on Overleaf.
.

--- 2019-02-05 main.tex
@@
-\input{header.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Standard header for working papers
+%
+% WPHeader.tex
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[11pt]{article}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% TEMPLATES
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Simple Tabular
+
+%\begin{tabular}{ |c|c|c| } 
+% \hline
+% cell1 & cell2 & cell3 \\ 
+% cell4 & cell5 & cell6 \\ 
+% cell7 & cell8 & cell9 \\ 
+% \hline
+%\end{tabular}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Packages
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% encoding 
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
+% general packages without options
+\usepackage{amsmath,amssymb,amsthm,bbm}
+
+% graphics
+\usepackage{graphicx,transparent,eso-pic}
+
+% text formatting
+\usepackage[document]{ragged2e}
+\usepackage{pagecolor,color}
+%\usepackage{ulem}
+\usepackage{soul}
+
+% conditions
+\usepackage{ifthen}
+
+\usepackage{booktabs}
+\usepackage{threeparttable}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Maths environment
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\newtheorem{theorem}{Theorem}[section]
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newtheorem{proposition}[theorem]{Proposition}
+%\newtheorem{corollary}[theorem]{Corollary}
+
+%\newenvironment{proof}[1][Proof]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{definition}[1][Definition]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{example}[1][Example]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{remark}[1][Remark]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+
+%\newcommand{\qed}{\nobreak \ifvmode \relax \else
+%      \ifdim\lastskip<1.5em \hskip-\lastskip
+%      \hskip1.5em plus0em minus0.5em \fi \nobreak
+%      \vrule height0.75em width0.5em depth0.25em\fi}
+
+%% Commands
+
+\newcommand{\noun}[1]{\textsc{#1}}
+
+%% Math
+
+% Operators
+\DeclareMathOperator{\Cov}{Cov}
+\DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\E}{\mathbb{E}}
+\DeclareMathOperator{\Proba}{\mathbb{P}}
+
+\newcommand{\Covb}[2]{\ensuremath{\Cov\!\left[#1,#2\right]}}
+\newcommand{\Eb}[1]{\ensuremath{\E\!\left[#1\right]}}
+\newcommand{\Pb}[1]{\ensuremath{\Proba\!\left[#1\right]}}
+\newcommand{\Varb}[1]{\ensuremath{\Var\!\left[#1\right]}}
+
+% norm
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+
+% argmin
+\DeclareMathOperator*{\argmin}{\arg\!\min}
+
+% amsthm environments
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
+\newtheorem{assumption}{Assumption}
+
+% geometry
+\usepackage[margin=2cm]{geometry}
+
+% layout : use fancyhdr package
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+
+\makeatletter
+
+\renewcommand{\headrulewidth}{0.4pt}
+\renewcommand{\footrulewidth}{0.4pt}
+\fancyhead[RO,RE]{\textit{Working Paper}}
+\fancyhead[LO,LE]{CASA/ISC-PIF}
+\fancyfoot[RO,RE] {\thepage}
+\fancyfoot[LO,LE] {\noun{J. Raimbault}}
+\fancyfoot[CO,CE] {}
+
+\makeatother
+
 \begin{document}
@@
 Process: Migration flows & NA \textbf{check if origins can change} \\
-Process: Transportation congestion & Link queuing
+Process: Transportation congestion & Link queuing \\
 \bottomrule

### 2019-08-06 paper.tex
Update on Overleaf.
.

@@
 We construct a large corpus of papers, linked through backward citations.
-Our contribution relies on the following points: (i) the open dataset we make available can be of significant interest for researcher interested in urban modeling; (ii) we map a broad part of approaches to urban modeling, and more particularly clarify the relations between microsimulation, land-use transport modeling, and agent-nased modeling.
+Our contribution relies on the following points: (i) the open dataset we make available can be of significant interest for researcher interested in urban modeling; (ii) we map a broad part of approaches to urban modeling, and more particularly clarify the relations between microsimulation, land-use transport modeling, and agent-based modeling.
 % (i) sensitivity analysis - info to reconstruct discipline with such empirical data ?

### 2019-08-09 paper.tex
Update on Overleaf.

@@
 
+classif of urban models \cite{batty2016classifying}
+
 \subsection*{Citation network collection}
@@
 We finally remove manually two communities corresponding to medical science and economics of taxes papers, which appear to be connected (most probably through methodological papers) but are not relevant for our study.
-The final filtered network has a size of % TODO
+The final filtered network, on which we will work in the following, has a size of % TODO
   .
@@
     \centering
-    %\includegraphics{figures/}}}
-    \caption{}
+    \includegraphics[width=\linewidth]{figures/sensitivity_fullnw_hdmax100_subnws_modularity.png}
+    \caption{\csentence{Modularity of subnetworks as a function of horizontal depth.}}
     \label{fig:modularity-hdepth}
@@
     \centering
-    \includegraphics{figures/coveragesubnws.png}
-    \caption{Relative coverage of sub-networks for each initial request, as a function of horizontal depth $h_d$.}
+    \includegraphics[width=0.52\linewidth]{figures/coveragesubnws.png}
+    \includegraphics[width=0.38\linewidth]{figures/pareto_vcount-modularity.png}
+    \caption{(\textit{Left}) Relative coverage of sub-networks for each initial request, as a function of horizontal depth $h_d$; (\textit{Right}) Optimization of network properties.}
     \label{fig:coverage-hdepth}
@@
 
+%%%%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\linewidth]{figures/overlaps_relative.png}
+%    \caption{Overlaps between subnetworks, for different horizontal depth values.}
+%\end{figure}
+
 \subsection*{Reconstructing endogenous disciplines}
 
-% rq ; microsim does not appear explicitely in main title -> finally "in minority" in urban modeling ;  other terms (in case of same approach) and other approaches are preefered.
+% cluster sizes
+The distribution of community size is fat-tailed, while however not being power-law distributed (presence of two regimes)
 
-% cluster sizes
-The distribution of community size is fat-tailed in a power-law manner, as scientific disciplines generally are. % TODO ref on that ?
- We can thus limit our study to the most important communities.
+%% largest communities (>1% of the network)
+%[1] "35 ; Urban growth modeling ; 11.1269367053429"
+%[1] "11 ; Built environment ; 10.0326547627319"%
+%[1] "31 ; Urban modeling/Luti ; 9.25102480372403"
+%[1] "17 ; Transportation microsimulation ; 7.41679983325228"
+%[1] "36 ; Complexity ; 5.75279649829778"
+%[1] "2 ; Pedestrian microsimulation ; 5.00590564857917"
+%[1] "16 ; Geosimulation ; 5.00243173765025"
+%[1] "1 ; Evacuation modeling ; 4.90168832071146"
+%[1] "34 ; Accessibility ; 4.73494059612312"
+%[1] "25 ; Population microsimulation ; 4.70020148683388"
+%[1] "10 ; Energy ; 3.44959355242132"
+%[1] "22 ; Congestion ; 3.30021538247759"
+%[1] "4 ; Logistics simulation ; 3.09525463767109"
+%[1] "27 ; Route choice ; 2.66101577155562"
+%[1] "28 ; Transportation physics ; 2.4247898283888"%
+%[1] "18 ; Transportation economics ; 2.36920725352602"
+%[1] "24 ; Spatial interactions ; 2.29625512401862"
+%[1] "20 ; Urban metabolism ; 1.87591190161884"
+%[1] "9 ; Critical infrastructures ; 1.62231640380741"
+%[1] "19 ; Social networks ; 1.60147293823386"
+%[1] "6 ; Demographics ; 1.43472521364552"
+%[1] "7 ; Mobility simulation ; 1.33050788577781"
+
+We describe in Table~\ref{tab:communities} the content of the largest communities (relative size larger than 2\% of network nodes).
+Several communities correspond to microsimulation approaches, but not most nor the largest.
+This is not due to an imbalance in subnetworks sizes, as the coverage graph in Fig.~\ref{fig:coverage-hdepth} already suggested.
+This confirms that urban modeling approaches considered are rather robust to the entry taken and that they form a broad consistent discipline.
+% largest is ca model - kind of microsim although they do not use the term
+
+%%%%%%%%%%%%%
+\begin{table}[h!]
+\caption{Description of main communities.\label{tab:communities}}
+% NOTE : for main papers, always taken among the two with largest in-degree [note done on purpose that so much Mike's papers !]
+      \begin{tabular}{|ccc|}
+        \hline
+        Community & Relative size (\%)  & Main papers   \\ \hline
+        Urban growth modeling & 11.1 & \cite{cheng2003urban} \cite{johnson2001environmental} \\
+        Built environment & 10.0 & \cite{ewing2010travel}  \\
+        Urban modeling/Luti & 9.3 & \cite{wegener2004overview} \cite{batty1976urban} \\
+        Transportation microsimulation & 7.4 & \cite{smith1995transims} \\
+        Complexity & 5.8 & \cite{batty2013new} \\
+        Pedestrian microsimulation & 5.0 & \cite{blue2001cellular} \\
+        Geosimulation & 5.0 & \cite{crooks2008key} \\
+        Evacuation modeling & 4.9 & \cite{cova2003network} \\
+        Accessibility & 4.7 & \cite{rodrigue2016geography} \\
+        Population microsimulation & 4.7 & \cite{beckman1996creating} \\
+        Energy & 3.4 & \cite{keirstead2012review} \\
+        Congestion & 3.3 & \cite{gomes2004congested} \\
+        Logistics microsimulation & 3.1 & \cite{hunt2007tour} \\
+        Route choice & 2.6 & \cite{ramming2001network} \\
+        Transportation physics & 2.4 & \cite{chowdhury2000statistical} \\
+        Transportation economics & 2.4 & \cite{small2007economics} \\
+        Spatial interactions & 2.3 & \cite{fotheringham1983new} \\\hline
+      \end{tabular}
+\end{table}
+%%%%%%%%%%%%%
 
@@
 We show in Fig.~\ref{fig:networkviz} a visualization of the network obtained with $h_d=3$.
+The relative positioning of communities is particularly interesting to understand proximities between subfields (we used a Force Atlas 2 network spatialization algorithm, which provide accurate information in terms of proximity although being not deterministic, since attraction forces betweem two communities will be shaped by the number of links between them).
 
-% Q : what about older nws ? also show interesting entry through one kw only. - we can not visualize everything unfortunately.
 %%%%%%%%
@@
     \centering
-    \includegraphics{figures/corehdepth3filteredLOWRES.png}
+    \includegraphics[width=\linewidth]{figures/corehdepth3filteredLOWRES.png}
     \caption{Network visualization for $h_d = 3$. Node color give community obtained with the Louvain algorithm.}
@@
 %Can we reconstruct the cluster based on iterative search terms in google scholar? -- see where the convergence point is for certain terms
+% . - index a large corpus with texts ; test different search algos ; investigate "semantic stability" of communities [cf sysalgoreview]
 

### 2019-08-12 paper.tex
.

@@
 % cluster sizes
-The distribution of community size is fat-tailed, while however not being power-law distributed (presence of two regimes)
+The distribution of community size is fat-tailed, while however not being power-law distributed (presence of two regimes, a first with communities with a size larger than 1\% of the network relatively balanced with a rank-size slope of -0.73 for an R-squared of 0.89, a second also relatively close to a power law with a fit of 0.76 but highly hierarchical with $\alpha = -1.5$ and negligible in any case).
+% TODO reformulate that shit
 
@@
 This confirms that urban modeling approaches considered are rather robust to the entry taken and that they form a broad consistent discipline.
+The largest community corresponds to urban cellular automata models, mostly applied to land-use change modeling and urban sprawl.
+This class of models is computational and very close to microsimulation indeed, but does not state so in semantic terms, what recalls the contingency of scientific descriptions and theories.
+We then find mostly expected (from an expert viewpoint % TODO Mike need your view on this
+) communities, including modeling the role on the built environment, urban modeling and land-use transport models, transport microsimulation models, complex urban science, and several either methodological or thematic domains (for example geosimulation and transportation economics are mostly disciplinary or methodological, while congestion, energy and logistics are defined by their object of study although they have been studied mostly by a single discipline.
 % largest is ca model - kind of microsim although they do not use the term
+% TODO IDEA : would it be possible to **endogenously** isolate these method/social/disci/thematic communities ?
+=> USING HIGHER ORDER INFO INCLUDING : coauthorshiups nws, semantic nws, cojournal / journal proximities , colab ; etc.
+==> multilayer approach could more effectively capture knowledge domains - multidimensional aspects of science.
+%The existence of clusters related to different tools or techniques (e.g. Matsim)
 
@@
 
-% Investigate the occurrence of some terms and their membership towards different clusters
-
-% graph visualization and relative positioning of disciplines
-
 We show in Fig.~\ref{fig:networkviz} a visualization of the network obtained with $h_d=3$.
 The relative positioning of communities is particularly interesting to understand proximities between subfields (we used a Force Atlas 2 network spatialization algorithm, which provide accurate information in terms of proximity although being not deterministic, since attraction forces betweem two communities will be shaped by the number of links between them).
+For this reason, it is natural to find at the core of the spatialized network the Luti, population and transportation microsimulation communities.
+% TODO check if complexity community is indeed merged with the CA one in the gephi community detection ; otherwise maybe change color as misleading.
+The complexity community is distributed as a glue between microsimulation approaches and urban modeling and Luti approaches, in accordance with the integrative ambition of these theories and models.
+Similarly, the transport geography (accessibility) community has its main papers at the interface between Luti and the built environment community.
+The network core is then surrounded by more or less peripheral communities, including statistical physics pedestrian models, urban energy models, logistics, transportation economics, congestion.
+% rq : in theory, the span/concentration of communities also means something : if they are highly self-citing, and is they are torn in different directions ( -> is this true indeed ? )
 
@@
     \includegraphics[width=\linewidth]{figures/corehdepth3filteredLOWRES.png}
-    \caption{Network visualization for $h_d = 3$. Node color give community obtained with the Louvain algorithm.}
+    \caption{\csentence{Core citation network visualization.} We show here the filtered core obtained for $h_d = 3$.
+Node color give community obtained with the Louvain algorithm.
+The network is spatialized using a Force atlas 2 algorithm, with the Gephi visualization software.}
     \label{fig:networkviz}
 \end{figure}
-%%%%%%%%
+%%%%%%%%s
+
+% proximity between communities (relative proportion of links)
+
+We compute the proximity between communities using a simple indicator, namely the proportion of out-citation links made outside the community.
+
+%%%%%%%%%%%
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{figures/com_intercit_proportion.png}
+	\caption{\csentence{Proximity matrix between communities.} We show the propostion of out-going links (note that the matrix is not symmetric as citation links are directed).}
+	\label{fig:outcitprop}
+\end{figure}
+%%%%%%%%%%%
 
@@
 
-%The existence of clusters related to different tools or techniques (e.g. Matsim)
+% -> distrib of interdisc index as diversity of citations (herfindhal on outcoming links \!/ check if significant as the network was not constructed this way [implies to be able to quantify this]
 
@@
 % with semantic content : 
+% Investigate the occurrence of some terms and their membership towards different clusters -> semantic ; as development ?
 %Text mining of abstracts to automatically denote -content- of the clusters\
